A prototype software framework for transferable computational health economic models

### 2024-02-20 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
moved URs

(i)	a prototype software framework we have developed to support the development of modular and open source CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->
<!-- (ii)  -->
(ii)	how we have applied that software framework to develop and share an initial set of modules for a economic model of youth mental health.

>>>>

(i)	our motivation for developing a transferable CHEM in youth mental health;

(ii) a prototype software framework we have developed to support the development of transferable CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->
<!-- (ii)  -->
(ii)	how we have applied that software framework to develop and share an initial set of modules for our youth mental health CHEM.

### 2024-02-20 Markdown/Manuscript/Child_RMDs/MS_Imp_Motivation.Rmd
moved URs

We became interested in modular and open-source approaches to CHEM development in 2018 when beginning to plan how to synthesise a number of current or planned economic modelling projects in youth mental health in one multi-purpose model. This overarching youth mental health model  is called readyforwhatsnext [@rfwn2024]. Initially, we plan to construct readyforwhatsnext from projects in four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:

@@

Although we are principally interested in using readyforwhatsnext to answer policy questions relating to the mental health of young people in Australia, we want to facilitate transferability by other teams to different jurisdictions. A high level schematic of how we aim to implement readyforwhatsnext as a modular and open source CHEM is outlined in Figure \@ref(fig:fig1).

@@

>>>>

We became interested in modular and open-source approaches to CHEM development in 2018 when beginning to plan how to synthesise a number of current or planned economic modelling projects in youth mental health in one multi-purpose model. This overarching youth mental health model is called readyforwhatsnext [@rfwn2024].

## Economic topics
Initially, we plan to construct readyforwhatsnext from projects in four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:

@@

A high level schematic of how we aim to implement readyforwhatsnext as a modular and open source CHEM to address these topics is outlined in Figure \@ref(fig:fig1).

@@

## User requirements
Although we are principally interested in using readyforwhatsnext to answer policy questions relating to the mental health of young people in Australia, we want to facilitate model transfer to other jurisdictions. We believe that a good starting point for facilitating model transferable is to develop a CHEM that is relatively easy to understand, use and adapt. For that reason, the initial user-requirements we specified for readyforwhatsnext are to meet six explicit criteria for transparency, reusability and updatability. We have describe these criteria elsewhere [@ethicsCHEMS2024], but briefly they are:

T1) Software files are open access;

T2) Developer contributions are easily identified;

T3) Programming practices promote selective reuse;

T4) Licenses restrict only unethical reuse; <!-- rephrase Liceses permit derivative works-->

T5) Maintenance infrastructure is in place; and

T6) New releases are systematically retested, with deprecation of outdated artefacts.

### 2024-02-20 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
moved URs

## Purpose
We decided to develop a software framework to help us address challenges we encountered during the planning and early implementation of readyforwhatsnext. The software framework is called ready4 [@ready424]. The purpose of ready4 is to support the development of CHEMs that meet explicit criteria for transparency, reusability and updatability. We have describe these criteria elsewhere [@ethicsCHEMS2024], but briefly they are: software files are open access, developer contributions are easily identified, programming practices promote selective reuse, licenses restrict only unethical reuse, maintenance infrastructure is in place and new releases are systematically retested, with deprecation of outdated artefacts.

@@

UR 3) A simple *syntax* that promotes support consistency in how algorithms associated with CHEM modules are described and used in health economic analysis programs.

@@

UR 6) *Reporting tools* that include: 
<!-- help generate tables for multiple outputs, retrieve templates, modify templates with authorship and results -->
(i) replicable health economic analysis programs; and (ii) reporting subroutines that.

UR 7) Tools to *find CHEM modules, datasets and replication programs*.

UR 8) Integration with tools for documenting open source software to help *develop and maintain a consolidated CHEM project website*.

>>>>

We decided to develop a software framework to help us address challenges we encountered during the planning and early implementation of readyforwhatsnext. The software framework is called ready4 [@ready424].

@@

UR 3) A *syntax* that promotes simplicity and consistency in how algorithms associated with CHEM modules are described and used in health economic analysis programs.

@@

UR 6) *Reporting tools* that: (i) help author programs for generating analysis reports and scientific summaries from a template that can render documents in HTML, Adobe PDF and Microsoft Word file formats; (ii) retrieve version specific reporting template programs from online code repositories and (iii) supply study specific data (e.g authorship details, input data and analysis results) to reporting template programs to render reports and scientific summaries.
<!-- help generate tables for multiple outputs, retrieve templates, modify templates with authorship and results -->

UR 7) *Search tools* to retrieve web based information on a CHEM project's: (i) publicly available model modules, module libraries and tutorials; (ii) datasets and data collections; and (iii) replication programs and reporting templates.

UR 8) *Documentation tools* that integrate with open source website development and hosting infrastructure to help: (i) develop a consolidated project documentation website that synthesises data from mulitple code library websites and code and data repositories; and (ii) partially automate the updating of website resources such as tutorials and code and data release information.

### 2024-02-20 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
moved URs

<!-- Import_start -->Computational models, particularly those addressing economic topics, have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, health economic models can be skills and resource intensive to undertake.<!-- Import_end --> One suggestion for making health economic modelling projects more tractable is for health economists to re-use each other's models [@Arnold2010]. A potential strategy for achieving this goal is to make computational health economic models (CHEMs) both modular and open-source. A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular]. Modular approaches are currently being used to facilitate the development of complex computational models in disciplines such as biology [@pan2021modular], ecology [@barros2023empowering] and neuroscience [@frazier2022robust]. In health economics the related and enabling concept of reference models has been recommended [@Afzali2013], but peer reviewed studies describing modular health economic models remain relatively rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013].

>>>>

<!-- Import_start -->Computational models, particularly those addressing economic topics, have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, health economic models can be skills and resource intensive to undertake.<!-- Import_end --> One suggestion for making health economic modelling projects more tractable is for health economists to re-use each other's models [@Arnold2010].

<!-- Imported_from_Ethics-->Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. <!-- End_Imported_from_Ethics -->A potential strategy for achieving both these goals is to make computational health economic models (CHEMs) both modular and open-source. A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular]. Modular approaches are currently being used to facilitate the development of complex computational models in disciplines such as biology [@pan2021modular], ecology [@barros2023empowering] and neuroscience [@frazier2022robust]. In health economics the related and enabling concept of reference models has been recommended [@Afzali2013], but peer reviewed studies describing modular health economic models remain relatively rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013].

### 2024-02-20 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
one table to go
extracted ethics component
added motivation
moved URs

# Ethical responsibilities and assessment criteria
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

@@

# Application in youth mental health
```{r child="../Child_RMDs/MS_Imp_Example.Rmd", echo=FALSE}
```

>>>>

# Motivation
```{r child="../Child_RMDs/MS_Imp_Motivation.Rmd", echo=FALSE}
```

@@

# Initial application
```{r child="../Child_RMDs/MS_Imp_Example.Rmd", echo=FALSE}
```

### 2024-02-20 Markdown/Manuscript/Parent_PDF/Springer.Rmd
moved URs

title: A prototype software framework for transparent, reusable and updatable computational health economic models
titlerunning: Framework for ethical computational models

>>>>

title: A prototype software framework for transferable computational health economic models
titlerunning: Framework for transferable computational models
